---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![](https://www.r-pkg.org/badges/version/SPMD?color=green)](https://cran.r-project.org/package=SPMD)
[![](http://cranlogs.r-pkg.org/badges/grand-total/SPMD?color=blue)](https://cran.r-project.org/package=SPMD)
[![R-CMD-check](https://github.com/RobinDenz1/SPMD/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/RobinDenz1/SPMD/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/RobinDenz1/SPMD/graph/badge.svg)](https://app.codecov.io/gh/RobinDenz1/SPMD)
<!-- badges: end -->

# SPMD

Author: Robin Denz

## Description

`SPMD` is a small R Package implementing the ***S**ymmetric **P**air **M**atching **D**esign*. Briefly, it is a self-controlled method that allows estimation of the causal effect of a possible re-current binary exposure on a possible re-current event. It is similar in spirit to the famous self-controlled case series (SCCS) design in the sense that it is bi-directional, using both time before and after exposures as controls (irrespective of events). However, unlike the SCCS method, it not only automatically adjusts for time-invariant confounders, but also automatically adjusts for all time-effects, without requiring any assumptions on the functional form of these time effects.

It does so by creating pairs of individuals, in which each individual acts as a control for the other individual at their respective exposure time. It can be shown that under mild assumptions of multiplicative effects and no interactions, both individual-level effects and time effects cancel out perfectly. For more details, please consult the associated paper.

## Installation

This package is not finished. It is still under active development, so breaking changes may occur at any time without warning. The developmental version may be installed from github using the `remotes` R-Package:

```R
library(remotes)

remotes::install_github("RobinDenz1/SPMD")
```

## Bug Reports and Feature Requests

If you encounter any bugs or have any specific feature requests, please file an [Issue](https://github.com/RobinDenz1/SPMD/issues).

## Examples

Suppose we have data that looks like this:

```{r}
library(data.table)
library(SPMD)

set.seed(1234)
data <- sim_example_data(n=500)
head(data)
```

This is a simple simulated dataset. Here, `A` is the time-dependent exposure and `Y` is the outcome. The probability for both exposure and outcome are time-dependent, thus making time itself a sort of confounder. In addition, `X` is a time-fixed confounder. Suppose though, that `X` was unmeasured. If we want to estimate the causal effect of `A` on `Y` adjusting for both, we could use:

```{r}
spmd <- sym_pair_matching(Surv(start, stop, Y) ~ A, data=data,
                          id=".id", risk_period=40, pairs="all",
                          estimator="moments", bootstrap=TRUE)
summary(spmd)
```

The true effect is 2.5 (see `?sim_example_data`), so the estimate is actually very close.

## Citation

If you use this package, please cite the associated article:

PUT PAPER HERE

## License

© 2026 Robin Denz

The contents of this repository are distributed under the GNU General Public License. You can find the full text of this License in this github repository. Alternatively, see <http://www.gnu.org/licenses/>.
