---
title: "SPMD Simulation Results"
author: "Robin Denz"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, fig.width=7, fig.height=7)
```

```{r}
library(ggplot2)
library(data.table)
library(binom)
```

# Introduction

This script takes in the results generated using the Monte-Carlo simulation and analyzes them, creating the plots and tables shown in the paper and its appendices. To reproduce these results, first execute the `run_simulation.r` script. Only then this markdown document may be created.

```{r}
# some general data preparations for simulation results
prepare_sim_results <- function(d_sim) {
  
  d_sim <- within(d_sim, {
    scenario <- paste0("Scenario ", scenario);
    estimator <- factor(
      estimator,
      levels = c("cox", "cco", "ctc", "sccs", "sccs_spline_5",
                 "sccs_spline_15", "spmd"),
      labels = c("Cox", "CCO", "CTC", "SCCS", "SCCS\nSpline\n(5)",
                 "SCCS\nSpline\n(15)", "SPMD")
    );
    exp_theta <- exp(theta);
    exp_theta_f <- factor(
      exp_theta,
      levels = c("0.7", "1", "1.5", "2.5", "5"),
      labels = c(expression(exp(theta)~`=`~0.7),
                 expression(exp(theta)~`=`~1),
                 expression(exp(theta)~`=`~1.5),
                 expression(exp(theta)~`=`~2.5),
                 expression(exp(theta)~`=`~5))
    );
    beta_L_A_f <- factor(
      exp(beta_L_A),
      levels = c("1", "3", "5"),
      labels = c(expression(exp(beta[A])~`=`~1),
                 expression(exp(beta[A])~`=`~3),
                 expression(exp(beta[A])~`=`~5))
    );
    beta_L_Y_f <- factor(
      exp(beta_L_Y),
      levels = c("1", "3", "5"),
      labels = c(expression(exp(beta[Y])~`=`~1),
                 expression(exp(beta[Y])~`=`~3),
                 expression(exp(beta[Y])~`=`~5))
    );
    risk_period_f <- factor(
      risk_period,
      levels = c("10", "50", "100", "150", "300"),
      labels = c(expression(tau~`=`~10),
                 expression(tau~`=`~50),
                 expression(tau~`=`~100),
                 expression(tau~`=`~150),
                 expression(tau~`=`~300))
    );
    log_bias_i <- log(rr_hat) - theta;
    log_mse_i <- (log(rr_hat) - theta)^2;
  })
  
  d_sim <- as.data.table(d_sim)
  
  # event and other counts were only made for estimator="spmd", so fill the
  # rest of the rows with the same (equivalent) info
  d_sim[, n_events := min(n_events, na.rm=TRUE),
        by=c("rep_id", "scenario", "theta", "n", "multiple_A", "multiple_Y",
             "beta_L_Y", "beta_L_A", "U_time_interact", "A_time_interact",
             "conf_int", "risk_period")]
  d_sim[, n_exposed_and_event := min(n_exposed_and_event, na.rm=TRUE),
        by=c("rep_id", "scenario", "theta", "n", "multiple_A", "multiple_Y",
             "beta_L_Y", "beta_L_A", "U_time_interact", "A_time_interact",
             "conf_int", "risk_period")]
  
  # remove rows where rr = 0
  d_sim <- subset(d_sim, !(rr_hat==0))
  
  return(d_sim)
}

# read in simulation results
d_sim1 <- readRDS("./data/sim1_results.Rds")
d_sim2 <- readRDS("./data/sim2_results.Rds")
d_sim3 <- readRDS("./data/sim3_results.Rds")
d_sim4 <- readRDS("./data/sim4_results.Rds")
d_sim5 <- readRDS("./data/sim5_results.Rds")
d_sim6 <- readRDS("./data/sim6_results.Rds")
d_sim7 <- readRDS("./data/sim7_results.Rds")
d_sim8 <- readRDS("./data/sim8_results.Rds")
d_sim9 <- readRDS("./data/sim9_results.Rds")

# prepare it for plotting
d_sim1 <- prepare_sim_results(d_sim1)
d_sim2 <- prepare_sim_results(d_sim2)
d_sim3 <- prepare_sim_results(d_sim3)
d_sim4 <- prepare_sim_results(d_sim4)
d_sim5 <- prepare_sim_results(d_sim5)
d_sim6 <- prepare_sim_results(d_sim6)
d_sim7 <- prepare_sim_results(d_sim7)
d_sim8 <- prepare_sim_results(d_sim8)
d_sim9 <- prepare_sim_results(d_sim9)
```

# Varying $n$ and $\theta$

```{r}
## simple boxplots for each estimator and scenario
plot_boxplots_n <- function(data, outliers=FALSE) {
  ggplot(data=data, aes(y=log_bias_i, x=estimator, fill=estimator)) +
    geom_hline(yintercept=0, linetype="dashed") +
    geom_boxplot(outliers=outliers, linewidth=0.3) +
    facet_grid(cols=vars(scenario), rows=vars(exp_theta_f),
               labeller = labeller(
                 exp_theta_f = label_parsed,
                 scenario = label_value
               )) +
    theme_minimal() +
    theme(legend.position="none",
          axis.text.x=element_text(size=8)) +
    labs(x=NULL, y=expression(hat(theta) - theta))
}

## create a table of bias, MSE and associated MCSE by n / theta / scenario
get_table_n_theta <- function(d_sim) {
  
  d_bias <- copy(as.data.table(d_sim))
  d_bias <- subset(d_bias, is.finite(rr_hat) & !is.na(rr_hat) & rr_hat!=0)

  d_bias <- d_bias[,
    .(bias = mean(log(rr_hat) - theta, na.rm=TRUE),
      bias_mcse = sd(log(rr_hat) - theta, na.rm=TRUE) / sqrt(1000),
      mse = mean((log(rr_hat) - theta)^2, na.rm=TRUE),
      mse_mcse = sd((log(rr_hat) - theta)^2, na.rm=TRUE) / sqrt(1000),
      n_NA = 1000 - .N,
      n_events = mean(n_events),
      n_events_exp = mean(n_exposed_and_event)),
    by=c("n", "exp_theta", "scenario", "estimator")
  ]

  setkey(d_bias, n, exp_theta, scenario, estimator)
  return(d_bias)
}
```

## $n = 5000$

```{r, results="asis"}
d_bias1 <- get_table_n_theta(d_sim1)
knitr::kable(subset(d_bias1, n==5000)[, -"n"], format="html", booktabs=TRUE,
             digits=3, linesep="", longtable=TRUE)
```

Boxplots not shown due to convergence issues in many methods.

## $n = 10000$

```{r}
plot_boxplots_n(subset(d_sim1, n==10000))
ggsave("./Plots/sim1_n10000.pdf", width=6.5, height=7)
```

```{r}
knitr::kable(subset(d_bias1, n==10000)[, -"n"], format="html", booktabs=TRUE,
             digits=3, linesep="", longtable=TRUE)
```

## $n = 20000$

```{r}
plot_boxplots_n(subset(d_sim1, n==20000))
ggsave("./Plots/sim1_n20000.pdf", width=6.5, height=7)
```

```{r}
knitr::kable(subset(d_bias1, n==20000)[, -"n"], format="html", booktabs=TRUE,
             digits=3, linesep="", longtable=TRUE)
```

## Runtime

```{r}
ggplot(subset(d_sim1, estimator=="SPMD"),
       aes(x=factor(n), y=runtime, fill=factor(exp_theta))) +
  geom_boxplot(size=0.1, linewidth=0.2) +
  facet_wrap(~ scenario, ncol=1) +
  theme_minimal() +
  theme(legend.position="bottom") +
  labs(x="n", y="Runtime in seconds", fill=expression(exp(theta)))
ggsave("./Plots/sim1_runtime.pdf", width=5, height=6)
```

# Multiple exposure periods per person

## $n = 10000$

```{r}
plot_boxplots_n(subset(d_sim2, n==10000))
ggsave("./Plots/sim2_n10000.pdf", width=6, height=6)
```

```{r}
d_bias2 <- get_table_n_theta(d_sim2)
knitr::kable(subset(d_bias2, n==10000)[, -"n"], format="html", booktabs=TRUE,
             digits=3, linesep="", longtable=TRUE)
```

## $n = 20000$

```{r}
plot_boxplots_n(subset(d_sim2, n==20000))
ggsave("./Plots/sim2_n20000.pdf", width=6, height=6)
```

```{r}
knitr::kable(subset(d_bias2, n==20000)[, -"n"], format="html", booktabs=TRUE,
             digits=3, linesep="", longtable=TRUE)
```

# Varying $n$ and $\theta$ with a terminal event

## $n = 10000$

```{r}
plot_boxplots_n(subset(d_sim3, n==10000))
ggsave("./Plots/sim3_n10000.pdf", width=6, height=6)
```

```{r}
d_bias3 <- get_table_n_theta(d_sim3)
knitr::kable(subset(d_bias3, n==10000)[, -"n"], format="html", booktabs=TRUE,
             digits=3, linesep="", longtable=TRUE)
```

## $n = 20000$

```{r}
plot_boxplots_n(subset(d_sim3, n==20000))
ggsave("./Plots/sim3_n20000.pdf", width=6, height=6)
```

```{r}
knitr::kable(subset(d_bias3, n==20000)[, -"n"], format="html", booktabs=TRUE,
             digits=3, linesep="", longtable=TRUE)
```

## With time-varying variable $L_t$

```{r}
## simple boxplots for each estimator and scenario
plot_boxplots_beta_L <- function(data, outliers=FALSE) {
  ggplot(data=data, aes(y=log_bias_i, x=estimator, fill=estimator)) +
    geom_hline(yintercept=0, linetype="dashed") +
    geom_boxplot(outliers=outliers, linewidth=0.3) +
    facet_grid(cols=vars(beta_L_A_f), rows=vars(beta_L_Y_f),
               labeller = label_parsed) +
    theme_minimal() +
    theme(legend.position="none",
          axis.text.x=element_text(size=7)) +
    labs(x=NULL, y=expression(hat(theta) - theta))
}

## create a table of bias, MSE and associated MCSE by n / theta / scenario
get_table_n_beta <- function(d_sim) {
  
  d_bias <- copy(as.data.table(d_sim))
  d_bias <- subset(d_bias, is.finite(rr_hat) & !is.na(rr_hat))
  d_bias[, exp_beta_L_A := exp(beta_L_A)]
  d_bias[, exp_beta_L_Y := exp(beta_L_Y)]

  d_bias <- d_bias[,
    .(bias = mean(log(rr_hat) - theta, na.rm=TRUE),
      bias_mcse = sd(log(rr_hat) - theta, na.rm=TRUE) / sqrt(1000),
      mse = mean((log(rr_hat) - theta)^2, na.rm=TRUE),
      mse_mcse = sd((log(rr_hat) - theta)^2, na.rm=TRUE) / sqrt(1000),
      n_NA = 1000 - .N),
    by=c("n", "exp_beta_L_Y", "exp_beta_L_A", "scenario", "estimator")
  ]

  setkey(d_bias, n, exp_beta_L_Y, exp_beta_L_A, scenario, estimator)
  return(d_bias)
}
```

## Scenario 1

```{r}
plot_boxplots_beta_L(subset(d_sim4, scenario=="Scenario 1"))
ggsave("./Plots/sim4_scenario1.pdf", width=7.5, height=7)
```

```{r}
d_bias4 <- get_table_n_beta(d_sim4)
knitr::kable(d_bias4[, -"n"], format="html", booktabs=TRUE,
             digits=3, linesep="", longtable=TRUE)
```

## Scenario 2

```{r}
plot_boxplots_beta_L(subset(d_sim4, scenario=="Scenario 2"))
ggsave("./Plots/sim4_scenario2.pdf", width=7.5, height=7)
```

```{r}
knitr::kable(d_bias4[, -"n"], format="html", booktabs=TRUE,
             digits=3, linesep="", longtable=TRUE)
```

# Bootstrap CI Coverage

```{r}
d_sim5[, in_ci := exp_theta >= ci_lower & exp_theta <= ci_upper]
d_sim5[, ci_length := ci_upper - ci_lower]

d_sim5_est <- d_sim5[, .(
  coverage = sum(in_ci) / .N,
  coverage_lower = binom.confint(x=sum(in_ci), n=.N, methods="exact")$lower,
  coverage_upper = binom.confint(x=sum(in_ci), n=.N, methods="exact")$upper,
  mean_length = mean(ci_length)
), by=c("n", "scenario", "exp_theta")  
]

ggplot(d_sim5_est, aes(x=exp_theta, y=coverage)) +
  geom_hline(yintercept=0.95, linetype="dashed") +
  geom_errorbar(aes(ymin=coverage_lower, ymax=coverage_upper), width=0.3) +
  geom_point() +
  facet_grid(cols=vars(scenario), rows=vars(n)) +
  theme_minimal() +
  scale_y_continuous(limits=c(0.85, 1), breaks=seq(0.85, 1, 0.05))
```

```{r}
knitr::kable(d_sim5_est)
```

# Interaction between time and $U$

## $n = 10000$

```{r}
plot_boxplots_n(subset(d_sim6, n==10000))
ggsave("./Plots/sim6_n10000.pdf", width=6, height=6)
```

## $n = 20000$

```{r}
plot_boxplots_n(subset(d_sim6, n==20000))
ggsave("./Plots/sim6_n20000.pdf", width=6, height=6)
```

# Interaction between time and $A$

## $n = 10000$

```{r}
plot_boxplots_n(subset(d_sim7, n==10000))
ggsave("./Plots/sim7_n10000.pdf", width=6, height=6)
```

## $n = 20000$

```{r}
plot_boxplots_n(subset(d_sim7, n==20000))
ggsave("./Plots/sim7_n20000.pdf", width=6, height=6)
```

# Different Baseline hazards per person

```{r}
ggplot(data=d_sim8, aes(y=log_bias_i, x=estimator, fill=estimator)) +
  geom_hline(yintercept=0, linetype="dashed") +
  geom_boxplot(outliers=FALSE, linewidth=0.3) +
  facet_wrap(~ exp_theta_f, labeller=label_parsed) +
  theme_minimal() +
  theme(legend.position="none",
        axis.text.x=element_text(size=8)) +
  labs(x=NULL, y=expression(hat(theta) - theta))
ggsave("./Plots/sim8.pdf", width=6, height=6)
```

# Varying $\tau$

```{r}
ggplot(data=subset(d_sim9, risk_period !=10), aes(y=log_bias_i, x=estimator, fill=estimator)) +
  geom_hline(yintercept=0, linetype="dashed") +
  geom_boxplot(outliers=FALSE, linewidth=0.3) +
  facet_grid(cols=vars(scenario), rows=vars(risk_period_f),
               labeller = labeller(
                 risk_period_f = label_parsed,
                 scenario = label_value
               )) +
  theme_minimal() +
  theme(legend.position="none",
        axis.text.x=element_text(size=8)) +
  labs(x=NULL, y=expression(hat(theta) - theta))
ggsave("./Plots/sim9.pdf", width=6, height=6)
```

