---
title: "SPMD Simulation Results"
author: "Robin Denz"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, fig.width=7, fig.height=7)
```

```{r}
library(ggplot2)
library(data.table)
library(binom)

# some general data preparations for simulation results
prepare_sim_results <- function(d_sim) {
  
  d_sim <- within(d_sim, {
    scenario_num <- scenario;
    scenario <- paste0("Scenario ", scenario);
    estimator <- factor(
      estimator,
      levels = c("cox", "cco", "ctc", "sccs", "sccs_spline_5",
                 "sccs_spline_15", "spmd"),
      labels = c("Cox", "CCO", "CTC", "SCCS", "SCCS\nSpline\n(5)",
                 "SCCS\nSpline\n(15)", "SPMD")
    );
    exp_theta <- exp(theta);
    exp_theta_f <- factor(
      exp_theta,
      levels = c("0.7", "1", "1.5", "2.5", "5"),
      labels = c(expression(exp(theta)~`=`~0.7),
                 expression(exp(theta)~`=`~1),
                 expression(exp(theta)~`=`~1.5),
                 expression(exp(theta)~`=`~2.5),
                 expression(exp(theta)~`=`~5))
    );
    beta_L_A_f <- factor(
      exp(beta_L_A),
      levels = c("1", "3", "5"),
      labels = c(expression(exp(beta[A])~`=`~1),
                 expression(exp(beta[A])~`=`~3),
                 expression(exp(beta[A])~`=`~5))
    );
    beta_L_Y_f <- factor(
      exp(beta_L_Y),
      levels = c("1", "3", "5"),
      labels = c(expression(exp(beta[Y])~`=`~1),
                 expression(exp(beta[Y])~`=`~3),
                 expression(exp(beta[Y])~`=`~5))
    );
    risk_period_f <- factor(
      risk_period,
      levels = c("10", "50", "100", "150", "300"),
      labels = c(expression(tau~`=`~10),
                 expression(tau~`=`~50),
                 expression(tau~`=`~100),
                 expression(tau~`=`~150),
                 expression(tau~`=`~300))
    );
    log_bias_i <- log(rr_hat) - theta;
    log_mse_i <- (log(rr_hat) - theta)^2;
    exp_beta_L_A <- exp(beta_L_A);
    exp_beta_L_Y <- exp(beta_L_Y);
  })
  
  d_sim <- as.data.table(d_sim)
  
  # event and other counts were only made for estimator="spmd", so fill the
  # rest of the rows with the same (equivalent) info
  d_sim[, n_events := min(n_events, na.rm=TRUE),
        by=c("rep_id", "scenario", "theta", "n", "multiple_A", "multiple_Y",
             "beta_L_Y", "beta_L_A", "U_time_interact", "A_time_interact",
             "conf_int", "risk_period")]
  d_sim[, n_exposed_and_event := min(n_exposed_and_event, na.rm=TRUE),
        by=c("rep_id", "scenario", "theta", "n", "multiple_A", "multiple_Y",
             "beta_L_Y", "beta_L_A", "U_time_interact", "A_time_interact",
             "conf_int", "risk_period")]
  
  # TODO: remove this after re-run of simulations
  if (!"allow_overlap" %in% colnames(d_sim)) {
    d_sim[, allow_overlap := FALSE]
    d_sim[, censor := 0]
  }
  
  # remove rows where rr = 0
  d_sim <- subset(d_sim, !(rr_hat==0))
  
  return(d_sim)
}

## simple boxplots for each estimator and scenario
plot_boxplots_n <- function(data, outliers=FALSE) {
  ggplot(data=data, aes(y=log_bias_i, x=estimator, fill=estimator)) +
    geom_hline(yintercept=0, linetype="dashed") +
    geom_boxplot(outliers=outliers, linewidth=0.3) +
    facet_grid(cols=vars(scenario), rows=vars(exp_theta_f),
               labeller = labeller(
                 exp_theta_f = label_parsed,
                 scenario = label_value
               )) +
    theme_minimal() +
    theme(legend.position="none",
          axis.text.x=element_text(size=8)) +
    labs(x=NULL, y=expression(hat(theta) - theta))
}

## simple boxplots for each estimator and scenario
plot_boxplots_beta_L <- function(data, outliers=FALSE) {
  ggplot(data=data, aes(y=log_bias_i, x=estimator, fill=estimator)) +
    geom_hline(yintercept=0, linetype="dashed") +
    geom_boxplot(outliers=outliers, linewidth=0.3) +
    facet_grid(cols=vars(beta_L_A_f), rows=vars(beta_L_Y_f),
               labeller = label_parsed) +
    theme_minimal() +
    theme(legend.position="none",
          axis.text.x=element_text(size=7)) +
    labs(x=NULL, y=expression(hat(theta) - theta))
}

## create a table of bias, MSE and associated MCSE by n / theta / scenario
get_stats_table <- function(d_sim, n_repeats=1000) {
  
  # remove infinite or NA values
  d_bias <- copy(as.data.table(d_sim))
  d_bias <- subset(d_bias, is.finite(rr_hat) & !is.na(rr_hat))
  
  # identify variables that were varied
  all_vars <- c("n", "exp_beta_L_Y", "exp_beta_L_A", "scenario_num", "estimator",
                "exp_theta", "multiple_A", "multiple_Y", "U_time_interact",
                "A_time_interact", "conf_int", "risk_period", "allow_overlap",
                "censor")
  varies <- apply(d_bias[, all_vars, with=FALSE], 2, function(x){uniqueN(x) > 1})
  all_vars <- all_vars[varies]
  
  # calculate performance stats
  d_bias <- d_bias[,
    .(bias = mean(log(rr_hat) - theta, na.rm=TRUE),
      bias_mcse = sd(log(rr_hat) - theta, na.rm=TRUE) / sqrt(.N),
      mse = mean((log(rr_hat) - theta)^2, na.rm=TRUE),
      mse_mcse = sd((log(rr_hat) - theta)^2, na.rm=TRUE) / sqrt(.N),
      n_events = mean(n_events),
      n_events_exp = mean(n_exposed_and_event),
      n_NA = n_repeats - .N),
    by=eval(all_vars)
  ]

  setkeyv(d_bias, all_vars)
  
  # put Bias / MSE and their MCSE in one columns
  d_bias[, `Bias [MCSE]` := paste0(round(bias, 3), " [", round(bias_mcse, 3), "]")]
  d_bias[, `MSE [MCSE]` := paste0(round(mse, 3), " [", round(mse_mcse, 3), "]")]
  d_bias[, c("bias", "bias_mcse", "mse", "mse_mcse") := NULL]
  
  # change column order
  setcolorder(d_bias,
              neworder=c(all_vars, "Bias [MCSE]", "MSE [MCSE]", "n_events",
                         "n_events_exp", "n_NA"))
  
  # change column names
  setnames(d_bias,
           old=c("exp_beta_L_Y", "exp_beta_L_A", "exp_theta", "risk_period",
                 "n_events", "n_events_exp", "n_NA", "scenario_num",
                 "estimator"),
           new=c("exp(beta_Y)", "exp(beta_A)", "exp(theta)",
                 "tau", "n_Y", "n_YA", "n NA",
                 "Scenario", "Estimator"),
           skip_absent=TRUE)
  
  return(d_bias)
}

## prints out the table of statistics
print_stats <- function(d_bias, format="html") {
  knitr::kable(d_bias, format=format, booktabs=TRUE, digits=3, linesep="",
               longtable=TRUE)
}
```

# Introduction

This script takes in the results generated using the Monte-Carlo simulation and analyzes them, creating the plots and tables shown in the paper and its appendices. To reproduce these results, first execute the `run_simulation.r` script. Only then this markdown document may be created.

```{r}
# read in simulation results
d_sim1 <- readRDS("./data/sim1_results.Rds")
d_sim2 <- readRDS("./data/sim2_results.Rds")
d_sim3 <- readRDS("./data/sim3_results.Rds")
d_sim4 <- readRDS("./data/sim4_results.Rds")
d_sim5 <- readRDS("./data/sim5_results.Rds")
d_sim6 <- readRDS("./data/sim6_results.Rds")
d_sim7 <- readRDS("./data/sim7_results.Rds")
d_sim8 <- readRDS("./data/sim8_results.Rds")
d_sim9 <- readRDS("./data/sim9_results.Rds")
d_sim10 <- readRDS("./data/sim10_results.Rds")
d_sim11 <- readRDS("./data/sim11_results.Rds")
d_sim12 <- readRDS("./data/sim12_results.Rds")
d_sim13 <- readRDS("./data/sim13_results.Rds")

# prepare it for plotting
d_sim1 <- prepare_sim_results(d_sim1)
d_sim2 <- prepare_sim_results(d_sim2)
d_sim3 <- prepare_sim_results(d_sim3)
d_sim4 <- prepare_sim_results(d_sim4)
d_sim5 <- prepare_sim_results(d_sim5)
d_sim6 <- prepare_sim_results(d_sim6)
d_sim7 <- prepare_sim_results(d_sim7)
d_sim8 <- prepare_sim_results(d_sim8)
d_sim9 <- prepare_sim_results(d_sim9)
d_sim10 <- prepare_sim_results(d_sim10)
d_sim11 <- prepare_sim_results(d_sim11)
d_sim12 <- prepare_sim_results(d_sim12)
d_sim13 <- prepare_sim_results(d_sim13)

# calculate bias tables for all scenarios (where it is needed)
d_bias1 <- get_stats_table(d_sim1)
d_bias2 <- get_stats_table(d_sim2)
d_bias3 <- get_stats_table(d_sim3)
d_bias4 <- get_stats_table(d_sim4)
d_bias6 <- get_stats_table(d_sim6)
d_bias7 <- get_stats_table(d_sim7)
d_bias8 <- get_stats_table(d_sim8)
d_bias9 <- get_stats_table(d_sim9)
d_bias10 <- get_stats_table(d_sim10)
d_bias11 <- get_stats_table(d_sim11)
d_bias12 <- get_stats_table(d_sim12)
d_bias13 <- get_stats_table(d_sim13)
```

# Varying $n$ and $\theta$

## $n = 5000$

Boxplots not shown due to convergence issues in many methods.

```{r}
print_stats(subset(d_bias1, n==5000)[, -"n"], format="html")
```

## $n = 10000$

```{r}
plot_boxplots_n(subset(d_sim1, n==10000))
ggsave("./Plots/sim1_n10000.pdf", width=6.5, height=7)
```

```{r}
print_stats(subset(d_bias1, n==10000)[, -"n"], format="html")
```

## $n = 20000$

```{r}
plot_boxplots_n(subset(d_sim1, n==20000))
ggsave("./Plots/sim1_n20000.pdf", width=6.5, height=7)
```

```{r}
print_stats(subset(d_bias1, n==20000)[, -"n"], format="html")
```

## Runtime

```{r}
ggplot(subset(d_sim1, estimator=="SPMD"),
       aes(x=factor(n), y=runtime, fill=factor(exp_theta))) +
  geom_boxplot(size=0.1, linewidth=0.2) +
  facet_wrap(~ scenario, ncol=1) +
  theme_minimal() +
  theme(legend.position="bottom") +
  labs(x="n", y="Runtime in seconds", fill=expression(exp(theta)))
ggsave("./Plots/sim1_runtime.pdf", width=5, height=6)
```

# Multiple exposure periods per person

## $n = 10000$

```{r}
plot_boxplots_n(subset(d_sim2, n==10000))
ggsave("./Plots/sim2_n10000.pdf", width=6, height=6)
```

```{r}
print_stats(subset(d_bias2, n==10000)[, -"n"], format="html")
```

## $n = 20000$

```{r}
plot_boxplots_n(subset(d_sim2, n==20000))
ggsave("./Plots/sim2_n20000.pdf", width=6, height=6)
```

```{r}
print_stats(subset(d_bias2, n==20000)[, -"n"], format="html")
```

# Varying $n$ and $\theta$ with a terminal event

## $n = 10000$

```{r}
plot_boxplots_n(subset(d_sim3, n==10000))
ggsave("./Plots/sim3_n10000.pdf", width=6, height=6)
```

```{r}
print_stats(subset(d_bias3, n==10000)[, -"n"], format="html")
```

## $n = 20000$

```{r}
plot_boxplots_n(subset(d_sim3, n==20000))
ggsave("./Plots/sim3_n20000.pdf", width=6, height=6)
```

```{r}
print_stats(subset(d_bias3, n==20000)[, -"n"], format="html")
```

## With time-varying variable $L_t$

## Scenario 1

```{r}
plot_boxplots_beta_L(subset(d_sim4, scenario=="Scenario 1"))
ggsave("./plots/sim4_scenario1.pdf", width=7.5, height=7)
```

```{r}
print_stats(subset(d_bias4, Scenario==1)[, -"Scenario"], format="html")
```

## Scenario 2

```{r}
plot_boxplots_beta_L(subset(d_sim4, scenario=="Scenario 2"))
ggsave("./plots/sim4_scenario2.pdf", width=7.5, height=7)
```

```{r}
print_stats(subset(d_bias4, Scenario==2)[, -"Scenario"], format="html")
```

# Bootstrap CI Coverage

```{r}
d_sim5[, in_ci := exp_theta >= ci_lower & exp_theta <= ci_upper]
d_sim5[, ci_length := ci_upper - ci_lower]

d_sim5_est <- d_sim5[, .(
  coverage = sum(in_ci) / .N,
  coverage_lower = binom.confint(x=sum(in_ci), n=.N, methods="exact")$lower,
  coverage_upper = binom.confint(x=sum(in_ci), n=.N, methods="exact")$upper,
  mean_length = mean(ci_length)
), by=c("n", "scenario", "exp_theta")  
]

ggplot(d_sim5_est, aes(x=factor(exp_theta), y=coverage)) +
  geom_hline(yintercept=0.95, linetype="dashed") +
  geom_errorbar(aes(ymin=coverage_lower, ymax=coverage_upper), width=0.3) +
  geom_point() +
  facet_grid(cols=vars(scenario), rows=vars(n)) +
  theme_minimal() +
  scale_y_continuous(limits=c(0.85, 1), breaks=seq(0.85, 1, 0.05))
ggsave("./plots/sim5.pdf", width=6, height=7)
```

```{r}
knitr::kable(d_sim5_est)
```

# Interaction between time and $U$

## $n = 10000$

```{r}
plot_boxplots_n(subset(d_sim6, n==10000))
ggsave("./Plots/sim6_n10000.pdf", width=6, height=6)
```

```{r}
print_stats(subset(d_bias6, n==10000)[, -"n"], format="html")
```

## $n = 20000$

```{r}
plot_boxplots_n(subset(d_sim6, n==20000))
ggsave("./Plots/sim6_n20000.pdf", width=6, height=6)
```

```{r}
print_stats(subset(d_bias6, n==20000)[, -"n"], format="html")
```

# Interaction between time and $A$

## $n = 10000$

```{r}
plot_boxplots_n(subset(d_sim7, n==10000))
ggsave("./Plots/sim7_n10000.pdf", width=6, height=6)
```

```{r}
print_stats(subset(d_bias7, n==10000)[, -"n"], format="html")
```

## $n = 20000$

```{r}
plot_boxplots_n(subset(d_sim7, n==20000))
ggsave("./Plots/sim7_n20000.pdf", width=6, height=6)
```

```{r}
print_stats(subset(d_bias7, n==20000)[, -"n"], format="html")
```

# With independent right-censoring

```{r}
plot_boxplots_n(d_sim11)
ggsave("./Plots/sim11.pdf", width=6, height=6)
```

```{r}
print_stats(d_sim11, format="html")
```

# With right-censoring dependent on $U$

```{r}
plot_boxplots_n(d_sim12)
ggsave("./Plots/sim12.pdf", width=6, height=6)
```

```{r}
print_stats(d_sim12, format="html")
```

# With right-censoring dependent on $L(t)$

## Scenario 1

```{r}
plot_boxplots_beta_L(subset(d_sim13, scenario=="Scenario 1"))
ggsave("./plots/sim13_scenario1.pdf", width=7.5, height=7)
```

```{r}
print_stats(subset(d_sim13, Scenario==1)[, -"Scenario"], format="html")
```

## Scenario 2

```{r}
plot_boxplots_beta_L(subset(d_sim13, scenario=="Scenario 2"))
ggsave("./plots/sim13_scenario2.pdf", width=7.5, height=7)
```

```{r}
print_stats(subset(d_sim13, Scenario==2)[, -"Scenario"], format="html")
```

# Different Baseline hazards per person

## $n = 10000$

```{r}
plot_boxplots_n(subset(d_sim8, n==10000))
ggsave("./Plots/sim8_n10000.pdf", width=6, height=6)
```

```{r}
print_stats(subset(d_sim8, n==10000)[, -"n"], format="html")
```

## $n = 20000$

```{r}
plot_boxplots_n(subset(d_sim8, n==20000))
ggsave("./Plots/sim8_n20000.pdf", width=6, height=6)
```

```{r}
print_stats(subset(d_sim8, n==20000)[, -"n"], format="html")
```

# Varying $\tau$

## $n = 10000$

```{r}
ggplot(data=subset(d_sim9, n==10000),
       aes(y=log_bias_i, x=estimator, fill=estimator)) +
  geom_hline(yintercept=0, linetype="dashed") +
  geom_boxplot(outliers=FALSE, linewidth=0.3) +
  facet_grid(cols=vars(scenario), rows=vars(risk_period_f),
               labeller = labeller(
                 risk_period_f = label_parsed,
                 scenario = label_value
               )) +
  theme_minimal() +
  theme(legend.position="none",
        axis.text.x=element_text(size=8)) +
  labs(x=NULL, y=expression(hat(theta) - theta))
ggsave("./Plots/sim9_n10000.pdf", width=6, height=6)
```

```{r}
print_stats(subset(d_sim9, n==10000)[, -"n"], format="html")
```

## $n = 20000$

```{r}
ggplot(data=subset(d_sim9, n==20000),
       aes(y=log_bias_i, x=estimator, fill=estimator)) +
  geom_hline(yintercept=0, linetype="dashed") +
  geom_boxplot(outliers=FALSE, linewidth=0.3) +
  facet_grid(cols=vars(scenario), rows=vars(risk_period_f),
               labeller = labeller(
                 risk_period_f = label_parsed,
                 scenario = label_value
               )) +
  theme_minimal() +
  theme(legend.position="none",
        axis.text.x=element_text(size=8)) +
  labs(x=NULL, y=expression(hat(theta) - theta))
ggsave("./Plots/sim9_n20000.pdf", width=6, height=6)
```

```{r}
print_stats(subset(d_sim9, n==20000)[, -"n"], format="html")
```

# Allowing partial pairs

## $n = 10000$

```{r}
plot_boxplots_n(subset(d_sim10, n==10000))
ggsave("./Plots/sim10_n10000.pdf", width=6, height=6)
```

```{r}
print_stats(subset(d_sim10, n==10000)[, -"n"], format="html")
```

## $n = 20000$

```{r}
plot_boxplots_n(subset(d_sim10, n==20000))
ggsave("./Plots/sim10_n20000.pdf", width=6, height=6)
```

```{r}
print_stats(subset(d_sim10, n==20000)[, -"n"], format="html")
```
